---
title: 'Reproducible Research-Course Project 1 '
author: "Abdelmalek Mohamed"
---

##Assignment Instructions
1.Code for reading in the dataset and/or processing the data
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report


##1
##Code for reading in the dataset and/or processing the data
```{r, echo = TRUE} 
library(ggplot2)

activity <- read.csv("/Users/savvy2/Documents/Reproducible research/activity.csv")

activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)

summary(activity)
```
##2. mean total number of steps taken per day?
```{r, echo = TRUE} 
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", col = "red", ylim = c(0,30), breaks = seq(0,25000, by=2500))
```
##3. mean and median of the total number of steps taken per day:
```{r, echo = TRUE} 
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
```
##4.Time series plot of the average number of steps taken
```{r, echo = TRUE}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, type = "l", col="red", lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")
```
##5.The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```

##6.Code to describe and show a strategy for imputing missing data
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, echo = TRUE}
sum(is.na(activity$steps))
```
Devise a strategy for filling in all of the missing values in the dataset. 
```{r, echo = TRUE}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r, echo = TRUE}

activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```
##7Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
```{r, echo = TRUE}

hist(total_steps_imputed$daily_steps, col = "red", xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```
Mean of the total number of steps taken per day
```{r, echo = TRUE}
mean(total_steps_imputed$daily_steps)
```
median of the total number of steps taken per day:
```{r, echo = TRUE}
median(total_steps_imputed$daily_steps)
```
##8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo = TRUE}

activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```
# panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r, echo = TRUE}

activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```